---
title: "Admissions Data"
author: "Winnie Kadzo Yaa"
date: "Monday, July 06, 2015"
output: pdf_document
---

 SYNOPSIS
------------------

In this report we aim at not showing any relationship between the admissions count, the date's of birth,date of admissions and the date of discharge of the patients.I made a number of summaries and out of them I concluded  that there was a high number of patients turnout in the months of January followed by July and the least turnout was experienced in the months of September and October.
Also the summaries showed that the highest and lowest number of patients admitted were between the age group of 1-5yrs and 16-20yrs respectively.

Data
-----------------

This report makes use of the dataset"admissions_data.csv".The data was obtained from the admissions records between the year 2002-2015.
```{r, echo=FALSE ,warning=FALSE ,message=FALSE}
library(dplyr)
```

```{r,comment=NA}

admissions_data <- read.csv("~/AProject/admissions_data.csv")
admissions<-tbl_df(admissions_data)
admissions
```
To show the top 1000 admissions we use the function below.
```{r,results='hide',tidy=FALSE}
print(admissions,n=1000)
```

In order to work with the dates we first change the date format hence easily read in R.
```{r,,echo=FALSE,warning=FALSE,message=FALSE}
library(lubridate)
```

```{r}
admissions$date_of_birth2<- as.Date(admissions$date_of_birth,format = '%d/%m/%Y')

admissions$date_admn <- as.Date(admissions$date_admn,format = '%d/%m/%Y')
```

I obtained the patients age by taking the difference between the date of admissions and the date of birth then divided by 365.25 to obtain the years.
```{r}
admissions$age_yrs <- as.integer((admissions$date_admn- admissions$date_of_birth2)/365.25)
```

I created 5 categories of age groups ie."Under 1yr","1-5yrs","6-10yrs","11-15yrs" and "16-20yrs" 
```{r}
admissions$age_cat[admissions$age_yrs<1] <-0

admissions$age_cat[admissions$age_yrs>=1 & admissions$age_yrs<=5] <-1

admissions$age_cat[admissions$age_yrs>=6 & admissions$age_yrs<=10] <-2

admissions$age_cat[admissions$age_yrs>=11 & admissions$age_yrs<=15] <-3 

admissions$age_cat[admissions$age_yrs>=16 & admissions$age_yrs<=20] <-4


admissions$age_cat<-factor(admissions$age_cat,labels=c("Under 1yr","1-5yrs","6-10yrs",
                                                      "11-15yrs","16-20yrs"),ordered=TRUE) 


```

```{r,comment=NA}
table(admissions$age_cat)
```

I obtained a summary of the age in years(age_yrs) as follows:
```{r,echo=FALSE,warning=FALSE,message=FALSE}
admissions<-subset(admissions, !is.na(age_yrs))
```
```{r,comment=NA}
summary(admissions$age_yrs)

var(admissions$age_yrs)#variance

sd(admissions$age_yrs)#standard deviation

```
I used the month() to extract the months the patients were admitted from the date of  there admission.
```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(lubridate)
```
```{r,results='hide',tidy=FALSE}
admissions$mnth1 <- month(admissions$date_admn)#adding a new object(mnth1 for months(char))

admissions$mnth1 <- factor(admissions$mnth1,labels=c("Jan","Feb","Mar","Apr","May","Jun"
                                        ,"Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE)

```
The table below shows the count of all admissions by month.It is noted that the month of January(5489) has the highest number of admissions while September(4173) was the least number of patients admitted. 
```{r,comment=NA}
table(admissions$mnth1)
```
A graph of general admissions aganist months of the year.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(ggplot2)
```
```{r,echo=FALSE}
a<-ggplot(admissions, aes(factor(mnth1)))
a<-a +geom_bar()
a<-a + geom_bar(fill="blue", colour="pink")
a<-a+labs(x="MONTHS",y="ADMISSIONS COUNT")
a<-a+ggtitle("General Monthly Admissions")
a<-a+theme_grey()
a<-a+coord_flip()
a
```
The table below shows the count of all admissions by age. It is noted that age group"1-5yrs" had the highest number of admissions of 26382 while the least admitted age group was that of "16-20yrs".
```{r,comment=NA}
table(admissions$age_cat)

```
A graph of general admissions aganist age group.

```{r,echo=FALSE}
c<-ggplot(admissions[!is.na(admissions$age_cat),],aes(factor(age_cat)))
c<-c+geom_bar()
c<-c+geom_bar(fill="limegreen",colour="blue")
c<-c+labs(x="AGE GROUP",y="ADMISSIONS COUNT")
c<-c+ggtitle("Admissions according to Age Group")
c
```
From the above tables and graphs I came up with the following assumptions:
```{r}
```
  -Since there was a high admission in the month of January it is assumed that  it is due to the extreme temperatures experienced during that month hence causing skin infections especially to young children in this case ages 1-5yrs.
 -Also the admissions may be as a result of dusty winds which are a normal occurrence in this month. These weather conditions increases the chances of patients under the said age group risk of suffering from respiratory  diseases like asthma.

```{r}
```
  -It is also noted that from the months of May to July there was a slight increase in the number of admissions. I assumed that it is due to the cold and rainy weather experienced during these times of the year.The age group that had high admissions during these particular months are those between "1-5yrs" simply because their body immune system is still very weak hence they cannot cope with the low temperatures. Most of them stand a risk of getting respiratory infections like pneumonia. 
```{r}
```
   -The increased admissions may also be assumed to be because of the high risk that      children especially in this age group (1-5yrs) face on Malarial infection that is at its highest during the rainy seasons.
```{r}
```


Below are tables and plots of the MONTHLY ADMISSIONS count aganist AGE GROUPS for each month.I generated them using a for loop control structure.
```{r,echo=FALSE,comment=NA}
admissions$k <-month(admissions$date_admn)#introduction of a new object"k"(Int)

for (i in 1:12) {
  
  print(table(admissions$age_cat[admissions$k==i]))
  
  mnth1<-month.name[i]
  title<- paste(mnth1,"admission")
  
  c5<-ggplot(admissions[admissions$k==i &!is.na(admissions$age_cat),],aes(factor(age_cat))) 
  c5<-c5+geom_bar()
  c5<-c5+geom_bar(fill="brown",colour="yellow")
  c5<-c5+labs(x="AGE GROUP",y="ADMISSIONS COUNT")
  c5<-c5+ggtitle(title)
  c5<-c5+theme_bw()
  print(c5)
}
```


































